---
title: "Inferential Data Analysis"
author: "Zheng Wu"
date: "June 7, 2015"
output: pdf_document
---

###Load the ToothGrowth data and perform some basic exploratory data analyses.
```{r, message=FALSE}
library(datasets)
library(ggplot2)
library(gridExtra)
data <- ToothGrowth
plot1 <- ggplot(data, aes(x = len)) + geom_histogram() + facet_grid(.~supp) + labs(title = "Histogram of len by supp")
plot2 <- ggplot(data, aes(x = len)) + geom_histogram() + facet_grid(.~dose) + labs(title = "Histogram of len by dose")
grid.arrange(plot1, plot2)
```

###Provide a basic summary of the data.
```{r}
summary(data)
```
###Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose.

####By supp OJ vs VC
```{r}
lOJ <- data$len[data$supp == "OJ"]
lVC <- data$len[data$supp == "VC"]
t.test(lOJ,lVC, alternative = "greater")
```
Since the p value of the one sided t-test is smaller that the 0.05 alpha threshould, I would reject the null hypothesis that the true difference in means in the two groups is 0 in favor of the alternative hypothesis that the true difference in means is greater than 0.

####By dose 0.5 vs 1
```{r}
l0.5 <- data$len[data$dose == 0.5]
l1 <- data$len[data$dose == 1]
t.test(l0.5,l1, alternative = "less")
```
Since the p value of the one sided t-test is smaller that the 0.05 alpha threshould, I would reject the null hypothesis that the true difference in means in the two groups is 0 in favor of the alternative hypothesis that the true difference in means is less than 0.

####By dose 1 vs 2
```{r}
l2 <- data$len[data$dose == 2]
t.test(l1,l2, alternative = "less")
```
Since the p value of the one sided t-test is smaller that the 0.05 alpha threshould, I would reject the null hypothesis that the true difference in means in the two groups is 0 in favor of the alternative hypothesis that the true difference in means is less than 0.

####By dose 0.5 vs 2
```{r}
t.test(l0.5,l2, alternative = "less")
```
Since the p value of the one sided t-test is smaller that the 0.05 alpha threshould, I would reject the null hypothesis that the true difference in means in the two groups is 0 in favor of the alternative hypothesis that the true difference in means is less than 0.

###Assumptions needed for all the above conclusions:
1.Each of the two populations being compared should follow a normal distribution.
2.The data used to carry out the test should be sampled independently and randomly from the two populations being compared.